---
title: "Reproducible research Wk 2 Assignment"
author: "Marc LALY"
date: "12/03/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading and preprocessing the data

Loading data from <https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip>

```{r}
setwd("~/Rfolder/Data/Reproducible Research Wk 2 Assignement")

if (!file.exists("activity.csv")) {
  download.file("https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip",destfile = "repdata%2Fdata%2Factivity.zip")
  unzip("repdata%2Fdata%2Factivity.zip")
}

activity <- read.csv("activity.csv",sep=",",header=TRUE,na.strings = "NA")
str(activity)
```
Converting "Date" as date format :
```{r Date}
activity$date <- as.Date(activity$date)
str(activity)
```

## What is mean total number of steps taken per day?

1. Calculate the total number of steps taken per day
```{r}
sum(activity$steps,na.rm = TRUE)
```
2. Histogram of the total number of steps per day :
```{r}
hist(activity$steps,xlab = "Steps")
```


Mean of steps taken per day :
```{r}
mean(activity$steps,na.rm=TRUE)
```
And median :
```{r}
median(activity$steps,na.rm=TRUE)
```

## What is the average daily activity pattern?

1. Time series plot (i.e. ðšðš¢ðš™ðšŽ = "ðš•") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r}
#First get the average number of steps per 5 minutes interval across all days with tapply and store it in meanStepsInterval
meanStepsInterval <- tapply(activity$steps,activity$interval,mean,na.rm=TRUE)
#Plotting average number of steps per day by 5-minute interval
plot(unique(activity$interval),meanStepsInterval,type='l',xlab = "Interval",ylab='Averaged Steps')
```


2. 5-minute interval, on average across all the days in the dataset that contains the maximum number of steps :
```{r}
meanStepsInterval[meanStepsInterval==max(meanStepsInterval)]
```

## Imputing missing values

1. Counting NAs :
```{r}
# summary function gives us the number of missing value per columns
summary(activity)
```
As we can wee, we have 2304 NAs in the "Steps" recorded.

2. and 3. Strategy for filling NAs : I chose to replace NA by the average steps for the corresponding interval :
```{r}
activity2 <- activity
for (i in seq(length(activity2$steps))){ 
  if (is.na(activity2$steps[i])) {
          num <- as.character(activity2$interval[i])
          activity2$steps[i] <- meanStepsInterval[num]
  }
}
summary(activity2)
```
We can see that we don't have NAs anymore and that 3rd Quantile has increased.
4. Let's check it graphically :
```{r}
library(ggplot2)
# We need to merge both DF into one to make an histogram
activity$activity <- "activity1"
activity2$activity <- "activity2"
activities <- rbind(activity,activity2)
activities$activity <- as.factor(activities$activity)

ggplot(activities, aes(steps, fill = activity)) + geom_histogram(alpha = 0.5, position = 'identity')
```


As we can see, replacing NAs by the average steps has increased steps values significantly.

## Are there differences in activity patterns between weekdays and weekends?

1. Creating a new variable "day" that precise "weekday" and "weekend". We use activities with replaced NAs:
```{r}
for (j in seq(length(activity$steps))) {
        
        if (weekdays(activity2$date[j])=="Samedi" | weekdays(activity2$date[j])=="Dimanche") {
                activity2$day[j] <- "weekend"
        } 
        else {
                activity2$day[j] <- "weekday"
                }
}
#Checking how weekdays and weekends we get :
table(activity2$day)
```
2. Plot of the 5-minutes interval and the average number of steps by weekdays and weekends :

```{r}
weekDays <- subset(activity2,activity2$day=="weekday")
weekEnds <- subset(activity2,activity2$day=="weekend")
weekDaysMean <- tapply(weekDays$steps,weekDays$interval,mean)
weekEndsMean <- tapply(weekEnds$steps,weekEnds$interval,mean)

par(mfrow=c(2,1))
plot(unique(activity2$interval),weekDaysMean,type = "l",ylab = "Average steps weekdays",xlab="",main = "Weekday")
plot(unique(activity2$interval),weekEndsMean,type="l",ylab = "Average steps weekends",xlab="Interval",main="Weekend")
```


As we can see, there are more steps taken on average during weekends rather than during weekdays.
